Key Derivation From Noisy Sources With More Errors Than Entropy

### 2014-09-23 trunk/constantRate/ConstCompFuzzExt-reorg.tex
worked on the reusability section

\begin{abstract}
Fuzzy extractors convert a noisy source of entropy into a consistent uniformly-distributed key.  In the process of eliminating noise, they  lose some of the entropy of the original source---in the worst case, as much as the logarithm of the number of correctable error patterns. For many practical sources, the error tolerance necessary to achieve reliable key derivation is larger than the entropy of the source.  Unfortunately, it is impossible to provide security for all sources with this property.  Most known approaches for building fuzzy extractors work in the worst case and cannot be used when a source has \emph{more errors than entropy}.

@@

The situation with reusability is even worse: the only known construction of reusable fuzzy extractors \cite{Boyen2004} requires very particular relationships between $w_i$ values, which are unlikely to hold in any practical source. \bnote{negative results stated in our contributions. keep there?}

@@

We switch to computational security to obtain constructions with additional features. Our second construction provides reusability (against computationally bounded adversaries).  Like the first construction,
it requires that a constant fraction of symbols contribute fresh entropy. However,
the reusability we obtain is very strong:  security holds even if the multiple readings $w_i$ used in $\gen$ are \emph{arbitrarily correlated}, as long as each $w_i$ \emph{individually} satisfies the above conditions as a distribution. The switch to computational security is necessary here: as discussed in \secref{sec:reusable}, it is impossible to obtain such reusability information-theoretically for sources with more errors than entropy.
This construction requires that the fraction of errors is subconstant.

@@

Most known constructions of fuzzy extractors put sufficient information in $p$ to recover the original $w$ from a nearby $w'$ during $\rep$ (this procedure is called a \emph{secure sketch}), and then apply a randomness extractor to $w$ to get $r$.
Unfortunately, the current techniques for building secure sketches do not work for sources with more errors than entropy, because they lose at least $\log |B_t|$ bits of entropy regardless of the source (this loss is necessary when the source is uniform~\cite[Lemma C.1]{DBLP:journals/siamcomp/DodisORS08}). Moreover, relaxing the security definition to computational will not improve this approach~\cite[Corollary 3.8, Theorem 3.10]{fuller2013computational}. Thus, we take a different approach and do not attempt to recover $w$.

@@

Ideally, a fuzzy extractor allows a user to authenticate to multiple different services/devices.  That is, instead of creating a single pair $(r, p)$ from an initial reading $w$, we read the noisy source multiple times~($w_1,..., w_q$) and produce $(r_1, p_1),..., (r_q, p_q)$.  This problem is known as a reusable fuzzy extractor~\cite{Boyen2004}.  There are two primary parameters of the definition: 1) $q$, the reusability of the fuzzy extractor and 2) what class of correlation is allowed between readings $w_i$ and $w_j$.

Boyen considers two versions of the problem, first where the adversary sees $p_1,..., p_q$~(outsider security).  Second where the adversary controls some subset of the servers and sees key generation on arbitrary $p$~(insider security).
Boyen constructs a outsider reusable cryptographic fuzzy extractor for unbounded $q$ when the correlation between readings is limited to a transitive isometric permutation groups. Boyen transforms this construction to insider security using random oracles. In both constructions, the correlation between  $w_1,..., w_q$ is limited to a transitive isometric permutation group.  Unfortunately, this class of correlation is relatively weak and unrealistic.
%\paragraph{Previous Negative Results}

Boyen~\cite{Boyen2004} considers information-theoretic fuzzy extractors with unbounded reusability.
However, even for this relatively weak class of correlations, Boyen shows that reusability comes at a real cost~\cite[Theorem 11]{Boyen2004}.  If the correlation between readings contains a set that generates a transitive isometric permutation group, any secure sketch must lose entropy proportional to the number of errors corrected.\footnote{This result strengthens the bounds of~\cite[Appendix C]{DBLP:journals/siamcomp/DodisORS08}, who show this loss is necessary for uniform $W_1$.} This means achieving reusability for sources with more errors than entropy is impossible in this setting.
%\bnote{old stuff}
%Their construction is secure when the correlations between $w_1,.., w_q$ come from a subgroup of permutations used in the construction of the fuzzy extractor.  They then show information-theoretic fuzzy extractors secure with unbounded reuse can be transformed into their construction~\cite[Theorem 11]{Boyen2004}.  This shows that correlation between $w_1,..., w_q$ must be restricted to this limited class.  There is no reason to expect that noise between repeated readings comes from a well-behaved subgroup of permutations.  Their result can be interpreted as saying that reusability is impractical.
%
%However, this result is inherently information-theoretic, the adversary must exhaust all randomness of the fuzzy extractor and computes statistical distance over exponential size strings.  There are two  restrictions to avoid this result: 1) limit the number of reuses of the fuzzy extractor 2) provide security against computational adversaries.  The second restriction implies the adversary will only observe a polynomial number of invocations of the fuzzy extractor.

To overcome this result, we consider computationally bounded adversaries.
%We present a non-adaptive definition and note that our results extend to the adaptive definition presented by Boyen~\cite[Definition 7]{Boyen2004}.\bnote{We need to come back to justifying this or provide a proof that our scheme works in the adaptive setting.}
%\begin{definition}[Insider Reusable Fuzzy Extractors]
%\label{def:reusable fuzz}
%Let $\mathcal{W}$ be a family of distributions over $\mathcal{M}$.  Let $(\gen, \rep)$ be a $(\mathcal{M}, \mathcal{W}, \kappa, t)$-computational fuzzy extractor that is $(\epsilon, s)$-hard with error $\delta$.  Let $W_1,..., W_q$ be a joint distribution such that $\forall i, W_i\in\mathcal{W}$.  Let $(R_i, P_i)$ be the distribution formed by running $\gen$ on the distribution $W_i$.  We say that $(\gen, \rep)$ is $q$-reusable if for all $i$
%\begin{align*}
%\delta^{\mathcal{D}_{s}}(&(R_1, P_1), ..., (R_i, P_i),..., (R_q, P_q), \\
%&(R_1, P_1),..., (U_\kappa, P_i),..., (R_q, P_q)) < \epsilon.
%\end{align*}
%\end{definition}
We modify the definition of Boyen~\cite[Definition 7]{Boyen2004}.  We discuss the differences in \apref{sec:comparing reusability}.

@@

Finally, we allow the distinguishing adversary to see other keys.  This is strictly stronger than the definition of Boyen.  However, it is weaker than the notion of insider reusable fuzzy extractors where the adversary sees key generation on arbitrary $p$.  This is a mixture of reusability and robustness~\cite{dkrs2006} of the fuzzy extractor.  Our reusable construction~(\consref{cons:sampling}) is not insider secure.

>>>>

\begin{abstract}
\lnote{abstract needs work}
Fuzzy extractors convert a noisy source of entropy into a consistent uniformly-distributed key.  In the process of eliminating noise, they  lose some of the entropy of the original source---in the worst case, as much as the logarithm of the number of correctable error patterns. For many practical sources, the error tolerance necessary to achieve reliable key derivation is larger than the entropy of the source.  Unfortunately, it is impossible to provide security for all sources with this property.  Most known approaches for building fuzzy extractors work in the worst case and cannot be used when a source has \emph{more errors than entropy}.

@@

The situation with reusability is even worse: the only known construction of reusable fuzzy extractors \cite{Boyen2004} requires very particular relationships between $w_i$ values, which are unlikely to hold in any practical source.

@@

We switch to computational security to obtain constructions with additional features. Our second construction provides reusability (against computationally bounded adversaries).  Like the first construction,
it requires that a constant fraction of symbols contribute fresh entropy. However,
the reusability we obtain is very strong:  security holds even if the multiple readings $w_i$ used in $\gen$ are \emph{arbitrarily correlated}, as long as each $w_i$ \emph{individually} satisfies the above conditions as a distribution. 
This construction requires that the fraction of errors is subconstant.

@@

Most known constructions of fuzzy extractors put sufficient information in $p$ to recover the original $w$ from a nearby $w'$ during $\rep$ (this procedure is called a \emph{secure sketch}), and then apply a randomness extractor to $w$ to get $r$.
Unfortunately, the current techniques for building secure sketches do not work for sources with more errors than entropy, because they lose at least $\log |B_t|$ bits of entropy regardless of the source. Moreover, this loss is necessary when the source is uniform~\cite[Lemma C.1]{DBLP:journals/siamcomp/DodisORS08} or when reusability against a sufficiently rich class of correlations is desired~\cite[Theorem 11]{Boyen2004}.

Addtionally, relaxing the security definition to computational will not improve this approach~\cite[Corollary 3.8, Theorem 3.10]{fuller2013computational}. Thus, we take a different approach and do not attempt to recover $w$.

@@

An additional desirable feature of fuzzy extractors is reusability~\cite{Boyen2004}. Inuitively, it is the ability to support multiple independent enrollments of the same value, allowing users to reuse the same biometric or PUF, for example, with multiple noncommunicating providers. More precisely, the algorithm $\gen$ may be run multiple times on correlated readings $w_1,..., w_q$ of a given source. Each time, $\gen$ will produce a different pair of values $(r_1, p_1),..., (r_q, p_q)$. Security for each extracted string $r_i$ should hold even in the presence of all the helper strings $p_1, \dots, p_q$ (even though the reproduction procedure $\rep$ at the $i$th provider obtains only a single $w'_i$ close to $w_i$ and uses a single helper string $p_i$). Because the multiple providers may not trust each other, a stronger security feature (which we satisfy) ensures that each $r_i$ is secure even when all $r_j$ for $j\neq i$ are also given to the adversary.

 Our ability to construct reusable fuzzy extractors depends on the types of correlations allowed among $w_1, \dots, w_q$. Boyen \cite{Boyen2004} showed how to do so when each $w_i$ is a shift of $w_1$ by a value that is oblivious to the value of $w_1$ itself (formally, $w_i$ is a result of a transitive isometry applied to $w_1$). Boyen also showed that even for this weak class of correlations, any secure sketch must lose at least $\log |B_t|$ entropy~\cite[Theorem 11]{Boyen2004}.

\lnote{why do we need this outsider/insider distinction here? push everything in this paragraph to the discussion of the definition in the appendix}
Boyen considers two versions of the problem, first where the adversary sees $p_1,..., p_q$~(outsider security).  Second where the adversary controls some subset of the servers and sees key generation on arbitrary $p$~(insider security).  
Boyen constructs a outsider reusable cryptographic fuzzy extractor for unbounded $q$ when the correlation between readings is limited to a transitive isometric permutation groups. Boyen transforms this construction to insider security using random oracles. In both constructions, the correlation between  $w_1,..., w_q$ is limited to a transitive isometric permutation group.  Unfortunately, this class of correlation is relatively weak and unrealistic.
%\paragraph{Previous Negative Results}

%\bnote{old stuff}
%Their construction is secure when the correlations between $w_1,.., w_q$ come from a subgroup of permutations used in the construction of the fuzzy extractor.  They then show information-theoretic fuzzy extractors secure with unbounded reuse can be transformed into their construction~\cite[Theorem 11]{Boyen2004}.  This shows that correlation between $w_1,..., w_q$ must be restricted to this limited class.  There is no reason to expect that noise between repeated readings comes from a well-behaved subgroup of permutations.  Their result can be interpreted as saying that reusability is impractical.
%
%However, this result is inherently information-theoretic, the adversary must exhaust all randomness of the fuzzy extractor and computes statistical distance over exponential size strings.  There are two  restrictions to avoid this result: 1) limit the number of reuses of the fuzzy extractor 2) provide security against computational adversaries.  The second restriction implies the adversary will only observe a polynomial number of invocations of the fuzzy extractor.

%To overcome this result, we consider computationally bounded adversaries.
%We present a non-adaptive definition and note that our results extend to the adaptive definition presented by Boyen~\cite[Definition 7]{Boyen2004}.\bnote{We need to come back to justifying this or provide a proof that our scheme works in the adaptive setting.}
%\begin{definition}[Insider Reusable Fuzzy Extractors]
%\label{def:reusable fuzz}
%Let $\mathcal{W}$ be a family of distributions over $\mathcal{M}$.  Let $(\gen, \rep)$ be a $(\mathcal{M}, \mathcal{W}, \kappa, t)$-computational fuzzy extractor that is $(\epsilon, s)$-hard with error $\delta$.  Let $W_1,..., W_q$ be a joint distribution such that $\forall i, W_i\in\mathcal{W}$.  Let $(R_i, P_i)$ be the distribution formed by running $\gen$ on the distribution $W_i$.  We say that $(\gen, \rep)$ is $q$-reusable if for all $i$
%\begin{align*}
%\delta^{\mathcal{D}_{s}}(&(R_1, P_1), ..., (R_i, P_i),..., (R_q, P_q), \\
%&(R_1, P_1),..., (U_\kappa, P_i),..., (R_q, P_q)) < \epsilon.
%\end{align*}
%\end{definition}
We modify the definition of Boyen~\cite[Definition 7]{Boyen2004} for the computational setting.  We discuss the differences between our definition and the definitions in Boyen's work \apref{sec:comparing reusability}. \lnote{why do we call it ``outsider''?}

@@

Finally, we allow the distinguishing adversary to see other keys \lnote{we do? then why do we call our definition ``outsider''?}.  This is strictly stronger than the definition of Boyen \lnote{you mean than the outsider security definition of Boyen?}.  However, it is weaker than the notion of insider reusable fuzzy extractors where the adversary sees key generation on arbitrary $p$.  This \lnote{what is ``this''?} is a mixture of reusability and robustness~\cite{dkrs2006} of the fuzzy extractor.  Our reusable construction~(\consref{cons:sampling}) is not insider secure \lnote{according to Boyen's notion, right? Why not say ``not robust'' and/or clarify that $r_i$ can leak but $p_i$ cannot be modified.}.
